In this notebook, the afl player statistics data from the library 'fitzRoy' will
be extracted and filtered into a raw file


Load the data
```{r}
library(fitzRoy)
library(dplyr)
```
```{r}
# collect the first year 
player_stats <- fetch_player_stats(season = 2012, source = "AFL")
player_stats <- subset(player_stats, select = c(providerId, utcStartTime, compSeason.shortName, round.name, round.roundNumber, venue.name, home.team.club.name, away.team.club.name, player.player.position, player.player.player.playerId, player.player.player.givenName, player.player.player.surname, teamId, timeOnGroundPercentage, goals, behinds, kicks, handballs, 	disposals, marks,	bounces,	tackles,	contestedPossessions,	uncontestedPossessions,	totalPossessions,	inside50s,	marksInside50,	contestedMarks,	hitouts,	onePercenters,	disposalEfficiency,	clangers,	freesFor,	freesAgainst,	dreamTeamPoints,	rebound50s,	goalAssists,	goalAccuracy,	ratingPoints,	turnovers,	intercepts,	tacklesInside50, teamStatus, player.playerId, player.jumperNumber, team.name))
for (i in seq(2013, 2022, 1)) {
  temp_stats <- fetch_player_stats(season = i, source = "AFL")
  temp_stats <- subset(temp_stats, select = c(providerId, utcStartTime, compSeason.shortName, round.name, round.roundNumber, venue.name, home.team.club.name, away.team.club.name, player.player.position, player.player.player.playerId, player.player.player.givenName, player.player.player.surname, teamId, timeOnGroundPercentage, goals, behinds, kicks, handballs, 	disposals, marks,	bounces,	tackles,	contestedPossessions,	uncontestedPossessions,	totalPossessions,	inside50s,	marksInside50,	contestedMarks,	hitouts,	onePercenters,	disposalEfficiency,	clangers,	freesFor,	freesAgainst,	dreamTeamPoints,	rebound50s,	goalAssists,	goalAccuracy,	ratingPoints,	turnovers,	intercepts,	tacklesInside50, teamStatus, player.playerId, player.jumperNumber, team.name))
  player_stats <- union(player_stats,temp_stats)
}
write.csv(player_stats, file = '../data/raw/player_statistics.csv')
```
the afl player details dataset only has information from players 2013+, which is the optimal choice (contains infomation on player position) 


Next collect all draft history from 2013 to 2022
2013, 2019- have position data for some reason, whilst others do not. 
To solve this issue we will also take positional data from each game to work out general position,
with their heigh/weight then be used to determine role e.g Key_forward
```{r}
# collect the first year 
draft_history <- fetch_player_details_afl(2013,   team = NULL,  comp = "AFLM")
for (i in seq(2014, 2022, 1)) {
  temp_history <- fetch_player_details_afl(i,   team = NULL,  comp = "AFLM")
  if (!('position' %in% colnames(temp_history))) {
    temp_history['position'] <- 'UNKNOWN'
  }
  draft_history <- union(draft_history,temp_history)
}
```
Now we can save this dataset
```{r}
write.csv(draft_history, file = '../data/raw/draft_history.csv' )
```
Finally, we need to save the player position data so inputation on a 
players position can be performed. The idea is first identifying whether 
the player is either: Ruck, Forward, Midfielder or Defender.
Then afterwards we can use physical data (heigh, weight) to
determine their type

for this we will use fetch_lineup
```{r}
# collect the first year 
game_position <- fetch_lineup(season = 2013, comp = "AFLM")
for (i in seq(2014, 2022, 1)) {
  temp_position <- fetch_lineup(season = i, comp = "AFLM")
  if ('lateChanges' %in% colnames(temp_position)) {
    myvars <- names(temp_position) %in% c('lateChanges')
    temp_position <- temp_position[!myvars]
  }
  game_position <- union(game_position,temp_position)
}
```
Now we can save the game position data
```{r}
write.csv(game_position, file = '../data/raw/game_positions.csv')
```
Also, we need to take the afltables draft history as well to get number of games
```{r}
tables_draft <- fetch_player_details(source = "afltables")
write.csv(tables_draft, file = '../data/raw/afltables_draft.csv')
```

Now we have finished downloading the data, and can begin cleaning/ "PREPORCESSING'
